### 2021-07-14 doc/thesis/chapters/implementation.tex
gnuradio

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{GNU Radio}
  \subsection{Introduction}
  \subsection{Transmitter}
  \subsection{Receiver}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{GNU Radio}

This section describes the implementation in GNU Radio.
A transmitter, as well as a receiver are implemented using this software.
Both can be deployed on actual hardware, so a continuous transmitter or receiver is built in a real environment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Introduction}

GNU Radio is a free and open-source software development toolkit to implement a software-defined radio.
The implementation is done by connecting functional blocks in a block design.
This abstracts the inner workings of each functional block and thus provides an implementation approach on a very high level of abstraction.
GNU Radio supports interfaces to external RF hardware \cite{SoftwareGnuRadio}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{External Hardware}

In order to interface with the real world, two different hardware devices are used.
Both devices are supported by GNU Radio and are briefly described here.

\subsubsection{USRP b200mini}

The USRP b200mini is a device that can be used to create a software-defined radio.
It resembles the size of a business card and is developed by Ettus Research.
A signal range from 70 MHz to 6 GHz can be covered by the front end.
The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.

GNU Radio supports this device with the blocks \textit{UHD: USRP Sink} and \textit{UHD: USRP Source}.
Therefore, it can be used as both a receiver and a transmitter.

\subsubsection{RTL-SDR}

The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
The device delivers IQ data via the USB interface \cite{RTLSDR}.

Based on its hardware, the RTL-SDR can only be used as a receiver.

GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
Multiple blocks are connected to a block diagram, that is very similar to the one in Fig.\ref{fig:bd_matlab_transmitter}.

Again, an audio file is used as an input, which is modulated as an FM channel signal.
The entire multiplex signal is then modulated onto an actual carrier frequency of 99 MHz, that is within the FM radio broadcast band.

take care in using the FM band - follow regulations in \ref{}!.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System Level Integration for Verification}

system level view: can be used as a local transmitter for testing and verification.
sent signal is known, thus the receiver can be verified
